---
title: "Generic converter of plate- to column-data"
subtitle: "Use with discretion and QC the results!"
author: "Mario Niepel"
date: "December 2020"
version: "0.1"
output:
     html_document:
          toc: true
          toc_depth: 2
          theme: united
version: 0.1 
---

THIS SCRIPT IS OBSOLETE

```{r include=FALSE}

require("tidyverse")

```


Summary and requirements:

- script takes raw data in plate format coming off readers (such as HTM, CTG plate reader, maybe IncuCyte) in plate-based formats and converts it to column-based format ready to merge with meta-data
- raw data blocks have to be annotated in the same manner as meta-data plate-based blocks
- raw data blocks have to be complete and need to be fully numerical to allow downstream computation
- annotation has to contain a Plate_ID somewhere
- all blocks read in by this script have to be identical readouts so they can be appended




```{r designation of files and directory}

# Import of meta data and organization of folders and files

# user input of which meta file to load
# format has to be according to specifications laid out above
in_file <- c("20201215_JM_raw_data_1.csv", "20201215_JM_raw_data_1.csv")

# prefix for name of assembled complete file
out_file <- "20201215_JM_raw_data_assembled"

# input and output directories 
in_dir <- "./data/"
out_dir <- "./work/"

```



```{r read data and creat output dir}

# check if out_dirs exists; if not make it
if (!dir.exists(out_dir)) {dir.create(out_dir)}


#// script can currently only deal with one plate at a time
#// v02 properly automated to take multi-file input
wells <- read_csv("./data/20201215_JM_raw_data_1.csv", col_names = FALSE)
Plate_ID <- "20201215_JM_01"

 
wells <- read_csv("./data/20201215_JM_raw_data_2.csv", col_names = FALSE)
Plate_ID <- "20201215_JM_02"
 





```


```{r clean up to dataframe}

# clean up the raw data and convert into a dataframe

# identify positions of "variable"
variables <- data.frame(which(wells=="variable", arr.ind=TRUE))       # identify x/y of 'variables' 
variables <- variables[order(variables$row),]                         # sort variables table by row

well_level <- data.frame(Plate_ID = (rep(as.character(Plate_ID), 384)),
                         Plate_Row = (rep(LETTERS[seq(1:16)],24)),
                         Plate_Col = as.integer(trunc((seq(1:384)+15)/16)),
                         Well_No = seq(1:384))

# add column for Well_name (A01-P24)
# Well_Name is concatenation of Plate_Row and Plate_Col
# str_pad adds a leading character ("0") to width (2)

well_level <- well_level %>% 
     mutate(Well_Name = paste0(Plate_Row, as.factor(str_pad(Plate_Col, 2, pad = "0")))) %>%

# sort by well number
     arrange(Well_No)

```


```{r}

######################################################################
### could write this as sapply as well
### variables is a data frame
### turn into a list that contains x/y pairs
### pass on this list to sapply to slice out and replace
######################################################################


# join 384 row well_level info with readouts

end_row <- 0
for (i in 1:nrow(variables)) {                                   

# empty out dataframe
     slice <- data.frame()
          
# put row/col value of current plate map in row and col
     row <- variables[i,"row"]
     col <- variables[i,"col"]
  
     
# identify variable type, names, and attribute
     type <- wells[row, col+1]
     name <- wells[row+1, col]
     attribute <- wells[row+1, col+1]

# slice out plate based on location of 'variables'
     if  ( i==nrow(variables) ) ( end_row <- nrow(wells) ) else ( end_row <- variables[i+1,"row"]-1 )
     slice <- wells %>% slice((row+2):end_row)

# remove NA rows and NA columns
     slice <- slice[rowSums(is.na(slice)) != ncol(slice),] # remove cols of all NA
     slice <- slice[,colSums(is.na(slice)) != nrow(slice)] # remove rows of all NA     

     values <- unlist(slice, use.names = F)
 #    length(values)

# append column to well_level dataframe
     if (type=="numeric") { 
               well_level <- cbind(well_level, as.numeric(values))
          } else {
               well_level <- cbind(well_level, values)
     }
# add proper metadata heading to new df column 'Name_Attribute' e.g drug_id
     names(well_level)[ncol(well_level)] <- name
}

# # extract level variables
# well_level_names <- c(colnames(well_level[5:ncol(well_level)]))
# plate_level_names <- c(colnames(plate_level[]))
# exp_level_names <- c(colnames(exp_level[])) 

```

```{r}
write_tsv(well_level, paste0(out_dir, Plate_ID, "_raw.tsv"))


```

