---
title: "Generic converter of plate- to column-data"
subtitle: "Use with discretion and QC the results!"
author: "Mario Niepel"
date: "December 2020"
version: "0.2"
output:
     html_document:
          toc: true
          toc_depth: 2
          theme: united

---

```{r include=FALSE}

library(tidyverse)
library(readxl)

```

Summary and requirements:

- script takes raw data in plate format coming off readers (such as HTM, CTG plate reader, maybe IncuCyte) in plate-based formats and converts it to column-based format ready to merge with meta-data
     - separate parsers exist for IncuCyte and PlateReader
     - this one is generic and therefore requires annotation of the input data blocks
- raw data blocks have to be annotated in the same manner as meta-data plate-based blocks
     - blocks of four
     - 'readout' (coding variable) - 'numeric' 
     - '<measurement>'
- raw data blocks have to be complete and need to be fully numerical to allow downstream computation
- there has to be a single pair of cells desinating which is the readout in any file or sheet for
     - 'Plate_ID' - 'plate id'
     - data with equal Plate_ID will be appended by column
     - data with different Plate_ID will be appended by row
-> the script will fail if there are not identical measures for all Plate IDs




```{r designation of files and directory}

#// Import of meta data and organization of folders and files

#// user input of which meta file to load
#// format has to be according to specifications laid out above
in_file <- c("20201218 1007 4T1 CT26_raw.xlsx")

#// name of assembled complete file
out_file <- "20201218_JM_raw_assembled"

               # #// additional test case
               # in_file <- c("20201215_JM_raw_data_1.csv", "20201215_JM_raw_data_2.csv")
               # 
               # #// name of assembled complete file
               # out_file <- "20201215_JM_raw_assembled"

#// input and output directories 
in_dir <- "./data/"
out_dir <- "./work/"

```


```{r sanity checks}

#// read filetype of in_files
in_file_type <- unique(tolower(str_extract(in_file, "\\....$|\\.....$")))

#// make sure that there is a single file_type
if (length(in_file_type) > 1) stop("Can't process more than one filetype at once.")


#// make sure file_types are compatible with input
if (in_file_type %in% c(".tsv", ".txt", ".csv", ".xls", ".xlsx") == F) stop("Filetype not recognized.")

#// check if out_dirs exists; if not make it
if (!dir.exists(out_dir)) {dir.create(out_dir)}

```



```{r read data}
#// empty out ls_data
ls_data <- vector(mode = "list")

if (in_file_type %in% c(".tsv", ".txt")) {
     #// read tab delimited files
     ls_data <- lapply(paste0(in_dir, in_file), function(x) read_tsv(x, col_names = FALSE)) 
     #// name the list items by names of in_file
     names(ls_data) <- str_split(in_file, "\\....$|\\.....$", 1)
     } else if (in_file_type == c(".csv")) {
          #// read comma delimited files
          ls_data <- lapply(paste0(in_dir, in_file), function(x) read_csv(x, col_names = FALSE)) 
          #// name the list items by names of in_file
          names(ls_data) <- str_split(in_file, "\\....$|\\.....$", 1)
     } else { 
          #// read excel files
          #// loop through each in_file individually and read in all sheets
          for (i in in_file) {
               #// read sheet names and number of sheets
               sheet_names <- excel_sheets(paste0(in_dir, i)) 
               # read all sheets of input file
               temp <- lapply( sheet_names, function(x) { read_excel(paste0(in_dir, i), x, col_names = FALSE ) } )
               # name all sheets in list
               names(temp) <- sheet_names
               # append named input list to input list
               ls_data <- append(ls_data, temp) }
     }

rm(temp, i, sheet_names)

```




```{r parse plate layout to dataframe}

parse <- function(in_df) {

     #// remove all NA rows from input df
          in_df <- in_df[rowSums(is.na(in_df)) != ncol(in_df),] #// remove cols of all NA

     #// identify Plate_ID from input df
          Plate_ID <- in_df[[which(in_df == "Plate_ID", arr.ind = T)[1,1], which(in_df == "Plate_ID", arr.ind = T)[1,2] + 1]]

     #// generate ouput df with Plate_ID, Plate_Row, Plate_Col, Well_No
          out_df <- data.frame(Plate_ID = (rep(Plate_ID, 384)),
                               Plate_Row = (rep(LETTERS[seq(1:16)],24)),
                               Plate_Col = as.integer(trunc((seq(1:384) + 15)/16)),
                               Well_No = seq(1:384))

     #// add column for Well_name (A01-P24)
     #// Well_Name is concatenation of Plate_Row and Plate_Col
     #// str_pad adds a leading character ("0") to width (2)
          out_df <- out_df %>% 
          mutate(Well_ID = paste0(Plate_Row, as.factor(str_pad(Plate_Col, 2, pad = "0")))) %>%
     #// sort by well number
          arrange(Well_No)

     #// identify positions of "readout"
          readouts <- data.frame(which(in_df == "readout", arr.ind = TRUE))       

     #// sort variables table by row
          readouts <- readouts[order(readouts$row),]                         

     #// join 384 plate format raw data with out_df
          end_row <- 0
          for (i in 1:nrow(readouts)) {   
     
          #// empty out dataframe
               temp <- data.frame()
        
          #// put row/col value of current plate map in row and col
               row <- readouts[i,"row"]
               col <- readouts[i,"col"]

          #// identify variable type, names, and attribute
               type <- in_df[row, col + 1]
               name <- in_df[row + 1, col]

          #// slice out plate based on location of 'variables'
               if  ( i == nrow(readouts) ) (end_row <- nrow(in_df) ) else (end_row <- readouts[i + 1,"row"] - 1)
               temp <- in_df %>% slice((row + 2):end_row)

          #// remove NA rows and NA columns
               temp <- temp[rowSums(is.na(temp)) != ncol(temp),] #// remove cols of all NA
               temp <- temp[,colSums(is.na(temp)) != nrow(temp)] #// remove rows of all NA     
          
          #// append column to well_level dataframe
               if (type == "numeric") { 
                         out_df <- cbind(out_df, as.numeric(unlist(temp, use.names = F)))
                    } else {
                         out_df <- cbind(out_df, unlist(temp, use.names = F))
               }
          #// add proper metadata heading to new df column 'Name_Attribute' e.g drug_id
               colnames(out_df)[ncol(out_df)] <- name
     }
     return(out_df)
          
}


```




```{r parse ls_data}

#// call data parsing for all list items read into ls_data
ls_out <- lapply(ls_data, function(x) parse(x))

#// give each list item the appropriate name
names(ls_out) <- names(ls_data)

#// read out Plate_IDs for each list item
Plate_ID <- sapply(ls_out, function(x) { X <- unlist(x, use.names = F); x <- as.character(x[1,1]) } )
Plate_ID <- unname(Plate_ID)

#// create empty merge list
temp <- vector(mode="list", length = length(unique(Plate_ID)))

#// loop over unique Plate_IDs
for (i in 1:length(unique(Plate_ID))) {
     #// cycle through all list items with each unique Plate_ID and do full_join
     temp[[i]] <- reduce(ls_out[which( Plate_ID==unique(Plate_ID)[i] ) ], full_join)
     }

#// perform a full rbind over all items in merge_list
ls_out <- reduce(temp, rbind)


```




```{r assmble and write data}

write_tsv(ls_out, paste0(out_dir, out_file, ".tsv"))


```


###// Version History


#// version 0.2
- written for 20201218_JM
"20201218 1007 4T1 CT26 for Mario (1).xlsx"
- back compatible with "20201215_JM_raw_data_1/2.csv" once annotation was adjusted to conform to new script

Implemented:
- output to TSV (use this from now on for all outputs)
- enforcement of some kind of Plate_ID output
- detection of CSV, TSV, Excel for reading
- reading of multi-sheet Excels
- merge of same Plate_IDs but different measurements
- merge of different Plate_IDs but same measurements
--> No flexibility in demand for square data




#// Version 0.1
- written for 20201215_JM

in_file <- c("20201215_JM_raw_data_1.csv", "20201215_JM_raw_data_1.csv")
out_file <- "20201215_JM_raw_data_assembled"
