---
title: "Raw/Meta-Data Merge"
subtitle: "Use with discretion and QC the results!"
author: "Mario Niepel"
date: "December 2020"
version: "0.5"
output:
     html_document:
          toc: true
          toc_depth: 2
          theme: united 
---

# To Do

- generate code outline with chunk names and hash tags for indentations
- function to plot 96/384 well plate based on well_level_ids
     - lapply to run through list of well_level_IDs
     - print 1 graphical outline per 'is_unique' on a per-plate basis
- html output for summary
- set up QC plots as plot.ly (using ggplotly)

1) clean up and comment current code -- include (import/output steps into markdown)
2) input info in 'markdown' area and transmit to code
3) fortify meta-data assembly
4) fortify raw/meta merger
5) automate initial cleanup/plotting of merged data
6) function-ize current scripts -- clear spots for user input/edits
7) allow for multi-input of meta data
          - all columns must be identical
          - all Exp ID need to be identical
          - all Plate ID need to be unique
          - just attach via rbind
          



```{r loading libraries, include=FALSE}

library(tidyverse)
library(patchwork)
library(huxtable)
library(DT)

```




# Summary and requirements:


This script takes a CSV file from Excel or another similar program containing metadata of experiments and converts it semi-automatically to a long dataframe suitable to receive output from virtually any other instrument that puts the data on a per-plate basis either into a single table or a row/column format. This script only deals with the metadata. 'raw_data_input' deals with the experimental results. A third script 'meta_raw_merge' combines the two and performs some basic QC. The thought is that the output from that script is ready for deeper analysis, plotting of results, extracting data for other experiments, ...

## MetaData requirements:

The file contains two parts: Plate-level data first and Well-level data second. For now only 384 well plates are supported, but the script should be easily adjustable for 96 well plates or any other plate-based format. Spaces in any headers will be replaced with underscores

## Plate-level data:

- A table with columns for variables/attributes with one row for each physical plate used in the experiment.
- **One column needs to be labeled 'Experiment_ID' and must contain an identical ID in each row.** Best practices take the date of 'day0' of the experiment and append your initials. For an experiment I started on April 16th 2019 the ID would be "20190416_MN"
- **One column needs to be labeled 'Plate_ID' and must contain a unique ID in each row that should not overlap with any other experiments** in case results will letter be merged. Best practices are to simply take the experiment ID and append _01, _02, ...This way there will be no overlap. Each entry has to be a complete column and contain a header and one entry for each plate. Missing values will create ambiguities when addressing the data later and can not be tolerated


## Well-level data:

- each well-level treatment needs a complete 'block' of information made of two parts. Variable information for the type of treatment and qualitative or quantitative information about the treatment for each well
- the type of treatment has to be a block of four entries:

+-------------+------------+
|  variable   | type       |
+-------------+------------+
|  descriptor | attribute  |
+-------------+------------+

"variable" needs to be entered for the script to recognize a new column
type needs to be entered as "numeric" or "factor" to designate if there is a continuous variable or not
descriptor denotes a treatment class that makes sense (e.g. drug, cytokine, media, gRNA, cell_no, ...)
if a given descriptor has multiple attributes they are combined with the descriptors. For all descriptors the most basic attribute should be ID. Additional ones for a drug could be: concentration, the vehicle, and treatment time. This would give drug_ID, drug_conc, drug_vehicle, and drug_treattime. And each one of these variables would get their own block of row/column entries.

- the qualitative/quantitative block for each variable needs to be regular and complete (no empty cells in 16 rows x 24 columns). Empty cells can be filled either with NA, NULL, NaN, or whatever other label makes sense to you. The dataframe will contain these entries as it is being assembled, but they can easily be filtered out later. 

```{r designation of files and directory}

# Import of meta data and organization of folders and files

# user input of which meta file to load
# format has to be according to specifications laid out
meta_in_file <- "Bin_metadata"

# user input of which raw data files to load
raw_in_files <- c("CRI128-pSTAT1-tSTAT1.txt", "CRI128-MAR-PARP14.txt")

# name of assembled complete data file
out_data <- paste0(meta_in_file, "_complete")
out_qc <- paste0(meta_in_file, "_QC")

# input and output directories 
in_dir <- "./data/"
out_dir_data <- "./data/"
out_dir_plots <- "./data/"

```




```{r file input & QC, include=FALSE}

# check if out_dirs exists; if not make it
if (!dir.exists(out_dir_data)) {dir.create(out_dir_data)}
if (!dir.exists(out_dir_plots)) {dir.create(out_dir_plots)}

#read input_file
suppressMessages(meta <- read_csv(paste0(in_dir, meta_in_file, ".csv"), col_names = FALSE))

# remove all NA rows and columns
meta <- meta[rowSums(is.na(meta)) != ncol(meta),]
meta <- meta[,colSums(is.na(meta)) != nrow(meta)]

# add filler into empty cell in 1/1 if empty

# Validate the presence of Experiment_ID and Plate_ID
if (length(which(meta=="Experiment_ID")) == 0) {
     stop("Experiment_ID not found. Please fix input.")
}

if (length(which(meta=="Experiment_ID")) > 2) {
     stop("Experiment_ID definition is not unique. Please fix input.")
}

if (length(which(meta=="Plate_ID")) == 0) {
     stop("Plate_ID not found. Please fix input.")
}


```




```{r split plates/wells, echo=FALSE, warning=TRUE}

# identify entry with 'Experiment_ID'
row <- which(meta=="Experiment_ID", arr.ind = T)[1]
col <- which(meta=="Experiment_ID", arr.ind = T)[2]

# set name of Experiment_ID
Experiment_ID <- toString(meta[row+1,col])

# subset all rows of meta that contain "experiment ID" 
# %in% checks the resulting vectorized evaluation of "(meta[,col] == "Experiment_ID" | meta[,col] == Experiment_ID)" and tests if it is TRUE
plates <- meta[(meta[,col] == "Experiment_ID" | meta[,col] == Experiment_ID) %in% TRUE, ]

# subset all rows of meta with the logical inverse of plates
wells <- meta[!((meta[,col] == "Experiment_ID" | meta[,col] == Experiment_ID) %in% TRUE),]

```


```{r clean up and QC, echo=FALSE, warning=TRUE}
# clean up and sanity check for plates 

# remove NA columns as above
plates <- plates[,colSums(is.na(plates)) != nrow(plates)]

# check for presence of any NA in rows
if (sum(rowSums(is.na(plates))) > 0) {
     stop("Plate annotation is incomplete. Missing values can't be interpreted.")
}

# convert first row into names and remove that row
colnames(plates) <- as.character(unlist(plates[1,]))
plates <- plates[-1,]

# Set up Experiment_ID level and Plate_ID

# check that there is only one entry in Experiment_ID
if (n_distinct(plates$Experiment_ID)!=1) {
     stop("Experiment_ID is not identical across all plates.")
}

# set up experiment_level dataframe and remove from plates
exp_level <- data.frame(Experiment_ID=plates$Experiment_ID)
plates <- plates[ , !(names(plates)=="Experiment_ID")]

# check that there are nrow entries in Experiment_ID
if (n_distinct(plates$Plate_ID)!=nrow(plates)) {
     stop("Plate_ID is not unique across all plates.")
}

# set up plate_level dataframe and remove from plates
plate_level <- data.frame(Plate_ID=plates$Plate_ID) 
plates <- plates[ , !(names(plates)=="Plate_ID")]

```


```{r parse plate section with plate/well specs, echo=FALSE, warning=TRUE}

# number of unique vals per column
unique_vals <- unname( sapply(plates, function(x) length(unique(x))) )  
unique_cols <- unname( sapply(plates, function(x) length(unique(x))==1) ) 

# append all unique columns to exp_level
exp_level <- cbind(exp_level, plates[ ,unique_vals==1])

# append the opposite selection to plate_level
plate_level <- cbind(plate_level, plates[ ,unique_vals!=1])

# create and insert Plate_No column
plate_level$Plate_No <- as.integer(1:nrow(plates)) 
plate_level <- plate_level[ , c(1, ncol(plate_level), 2:(ncol(plate_level)-1))]

# Generate Plate_Row, _Col, Well_No
well_level <- data.frame(Plate_Row = (rep(LETTERS[seq(1:16)],24)),
                         Plate_Col = as.integer(trunc((seq(1:384)+15)/16)),
                         Well_No = seq(1:384))

# add column for Well_name (A01-P24)
well_level <- well_level %>% 
     # Well_Name is concatenation of Plate_Row and Plate_Col
     # str_pad adds a leading character ("0") to width (2)
     mutate(Well_Name = paste0(Plate_Row, as.factor(str_pad(Plate_Col, 2, pad = "0")))) %>%
     # sort by well number
     arrange(Well_No)

```


```{r extract well level annotation, echo=FALSE, warning=TRUE}

######################################################################
### could write this as mapply as well or lapply with string/split
### variables is a data frame
### turn into a list that contains x/y pairs
### pass on this list to sapply to slice out and replace
######################################################################

# identify positions of "variable"
variables <- data.frame(which(wells=="variable", arr.ind=TRUE))       # identify x/y of 'variables' 
variables <- variables[order(variables$row),]                         # sort variables table by row

# join 384 row well_level info with readouts
end_row <- 0
for (i in 1:nrow(variables)) {                                   

# empty out dataframe
     slice <- data.frame()
          
# put row/col value of current plate map in row and col
     row <- variables[i,"row"]
     col <- variables[i,"col"]

# identify variable type, names, and attribute
     type <- wells[row, col+1]
     name <- wells[row+1, col]
     attribute <- wells[row+1, col+1]

# slice out plate based on location of 'variables'
     if  ( i==nrow(variables) ) ( end_row <- nrow(wells) ) else ( end_row <- variables[i+1,"row"]-1 )
     slice <- wells %>% slice((row+2):end_row)

# remove NA rows and NA columns
     slice <- slice[rowSums(is.na(slice)) != ncol(slice),] # remove cols of all NA
     slice <- slice[,colSums(is.na(slice)) != nrow(slice)] # remove rows of all NA     

     values <- unlist(slice, use.names = F)
 #    length(values)

# append column to well_level dataframe
     if (type=="numeric") { 
               well_level <- cbind(well_level, as.numeric(values))
          } else {
               well_level <- cbind(well_level, values)
     }
# add proper metadata heading to new df column 'Name_Attribute' e.g drug_id
     names(well_level)[ncol(well_level)] <- paste0(name,"_",attribute)
}

# extract level variables
well_level_names <- c(colnames(well_level[5:ncol(well_level)]))
plate_level_names <- c(colnames(plate_level[]))
exp_level_names <- c(colnames(exp_level[])) 

```



```{r assemble plate and well data, echo=FALSE, warning=TRUE}

# assemble the full data frame from plate and well info

# join experiment and plate-level metadata
# Duplicate each row in plates 384x
# lastly reorder the file by plate No 
exp_plate_level <- cbind(exp_level, plate_level) %>% slice(rep(row_number(), 384)) %>% arrange(Plate_No)

# combine wells_long and plates_long
meta_processed <- cbind(exp_plate_level, well_level)

# sort by plate and well number
meta_processed <- arrange(meta_processed, Plate_No, Plate_Col)

# give proper row names to complete file
rownames(meta_processed) <- c(seq(1:nrow(meta_processed)))

# rearrange columns by using relocate
# go via creation of 'new order' vector in case exp_level or plate_level is empty
# this causes dplyr to throw an error
new_order <- c("Experiment_ID", "Plate_ID", "Plate_No", "Plate_Row", "Plate_Col", "Well_Name", "Well_No", exp_level_names, plate_level_names) 
meta_processed <- meta_processed %>% relocate (new_order[]) 

# save processed data
write_csv(meta_processed,paste0(out_dir_data, meta_in_file, "_formatted.csv"))


```






```{r data summary, echo=FALSE}

# table representation of experiment level annotation

# select only columns with names in 'exp_level_names'
exp_level_table <- meta_processed[1,exp_level_names]
# extract values w/o names as characters from every column
exp_level_values <- list(values=unlist(lapply(exp_level_table, as.character), use.names = FALSE))
# generate new df with name and value column
exp_level_table <- data.frame(names=exp_level_names, values = exp_level_values )
# experiment_level table
datatable(exp_level_table)

# table representation of plate level annotation

# select all columns based on plate level name and extract distinct values based on each Plate ID
plate_level_table <- meta_processed %>% select(all_of(plate_level_names)) %>% group_by(Plate_ID) %>% distinct()
# plate_level table
datatable(plate_level_table)

# table representation of well level annotation

# grepl("_ID", well_level) Searches for any list entry that contains the string "_ID" in the list of "well_level"
# substr cutting of the last three characters "_ID" through substring(x, 1, nchar(x)-3)
well_level_ids <- substr(well_level_names[grepl("_ID", well_level_names)], 1, nchar(well_level_names[grepl("_ID", well_level_names)])-3)
# subset all columns that start with well_level_ids
well_level_table <- meta_processed %>% select(all_of(starts_with(well_level_ids))) %>% lapply(unique)
# convert into characters so table can be better manipulated
well_level_table <- lapply(well_level_table, as.character)
# fill up empties with NA and turn into a dataframe
# lapply makes length of all vectors the same as the max which fills them up with NA
well_level_table <- data.frame(lapply(well_level_table, function(x) { length(x) <- max(lengths(well_level_table)); x }))
# well_level table
datatable(well_level_table)

# full metadata
datatable(meta_processed, options = list(
  columnDefs = list(list(className = 'dt-center', targets = 1:ncol(meta_processed))),
  pageLength = 16,
  lengthMenu = c(8, 16, 96, 384), class = "stripe hover compact",
  style="width:100%",
  scrollX = T)
)



```



```{r visual data summary, echo=FALSE, fig.width=16, fig.height=12 }

####################################################################################

qc_plot <- function(col, plate, df) { 

     nudge <- c( rep(0.25, times=16), rep(-0.25, times=16))    
    
     label <- col
     if (str_sub(label, start= -3) != "_ID" ) { label <- paste0(unlist(strsplit(label, "_"))[[1]], "_ID")}

     df %>% filter(Plate_ID == plate) %>%
     
     ggplot(aes(x = Plate_Col, y = Plate_Row)) + 
     coord_fixed() + scale_x_discrete(limits=c(as.character(seq(1, 24)))) +
     theme(legend.position="bottom", legend.box = "horizontal", text = element_text(size = 20)) +
     
     labs(title = paste("Readout:", col, "\n1st unique plate:", plate) ,
     caption = paste("Experiment ID:", Experiment_ID)) +
     
     geom_tile(aes(fill = .data[[col]]), color = "black", size = 0.5) +
     scale_fill_brewer(palette = "Blues", na.value = "gray") +
     geom_text(aes(x = Plate_Col, y = Plate_Row, label = .data[[label]]), check_overlap = TRUE, nudge_y = nudge, size = 5) 

}

####################################################################################

# generate lookup table of unique plates that need to be summarized
summary_plates <- 
     meta_processed %>%
     group_by(Plate_ID) %>%
     summarize(across(all_of(well_level_names), paste, collapse = ""), .groups = "drop") %>%
     mutate(across(all_of(well_level_names), ~!duplicated(.))) 

# generate list with entries for each colum listing plates to plot
list_summary_plates <- sapply(summary_plates[-1], function(x) { summary_plates$Plate_ID[x] })

# generate sum_plate and sum_ID vectors for summary plots
sum_IDs <- c(rep(names(list_summary_plates), lengths(list_summary_plates)))
sum_plates <- as.character(unlist(list_summary_plates, use.names = F))

# call qc_plot with mapply using the IDs and corresponding plate as list arguments
qc_plots <- mapply(qc_plot, sum_IDs, sum_plates, MoreArgs = list(meta_processed), SIMPLIFY = FALSE)

# print off all QC plots
# maybe set up as plot.ly (using ggplotly)
for (p in 1:length(qc_plots)) { print(qc_plots[[p]]) }

# Export to PDF
#   
# pdf(file = paste0(out_dir, out_file, ".pdf"),   # The directory you want to save the file in
#     width = 16, # The width of the plot in inches
#     height = 12) # The height of the plot in inches
# 
# for (i in 1:4) {
#      print(composite1[[i]])
#      print(composite2[[i]])
# }
# dev.off()
# invisible(NULL)

```





```{r raw data import and assembly, warning=TRUE, include=FALSE}

# grab meta from meta_processed
meta <- meta_processed

# read all data files into a list using sapply
raw <- lapply(paste0(in_dir, raw_in_files), function(x) read_tsv(x, col_names = TRUE))

# Plates in both files are the upper and lower half of the plate
# since they are independent scan they have different scan IDs
# they map to identical plate numbers
# harmonize Plate/Scan ID numbers for pSTAT1/tSTAT1
# data[[1]] gets the 1st element out of the list (instead of the packed list item)
# element gets written back after every row element in the first column gets eight added to match Plate ID in next file
raw[[1]][,1] <- raw[[1]][,1]+8

# reduce left iterates through a list with many items and consecutively applies the function to pairs in the list
# that's how reduce works to cycle through columns when checking for NA, but not through rows.
raw <- raw %>% reduce(full_join)

####################################################################################################
##### check here if there are multiple entries for the same well
##### use summary data function in dplyr to merge here? 
##### probably fine to use take 'sum or mean'
##### sum doesn't make sense if there is a 'fractional value'
##### mean doesn't make sense for a count
##### sum is probably the better way to go since it is easier to just recalculate fractional values
####################################################################################################

```


```{r clean up & merge with meta, warning=TRUE, include=FALSE}

# Clean up data to be merged

# Clean up column names
colnames(raw)[which(colnames(raw)=="% Positive W2 (MultiWaveScoring)")] <- "Fraction_W2"
colnames(raw)[which(colnames(raw)=="Cell: W2 Nucleus Average Intensity (MultiWaveScoring)")] <- "Nuc_W2"
colnames(raw)[which(colnames(raw)=="Cell: W2 Cytoplasm Average Intensity (MultiWaveScoring)")] <- "Cyto_W2"
colnames(raw)[which(colnames(raw)=="Cell: W2 Cell Average Intensity (MultiWaveScoring)")] <- "Cell_W2"
colnames(raw)[which(colnames(raw)=="% Positive W3 (MultiWaveScoring)")] <- "Fraction_W3"
colnames(raw)[which(colnames(raw)=="Cell: W3 Nucleus Average Intensity (MultiWaveScoring)")] <- "Nuc_W3"
colnames(raw)[which(colnames(raw)=="Cell: W3 Cytoplasm Average Intensity (MultiWaveScoring)")] <- "Cyto_W3"
colnames(raw)[which(colnames(raw)=="Cell: W3 Cell Average Intensity (MultiWaveScoring)")] <- "Cell_W3"
colnames(raw)[which(colnames(raw)=="Total Cells (MultiWaveScoring)")] <- "Cell_No"
colnames(raw) <- gsub(" ", "_", colnames(raw))

# remove all rows where Plate_ID is NA
raw <- raw[ !is.na(raw$Plate_ID)  , ]

# turn Plate_ID into chr
raw$Plate_ID <- as.character(raw$Plate_ID, stringsAsFactors = F)

# # set column to be used for merging raw data     
# raw_key <- as.character(unique(raw$Plate_ID))
# 
# # generate unique 'Join' column in data set from plate ID 
# raw <- cbind.data.frame(join_Plate_ID=as.character(raw$Plate_ID), raw[-1], stringsAsFactors = F)

# # generate unique 'Join' column in data set from plate ID 
# meta <- cbind.data.frame(join_Plate_ID=as.character(meta$Plate_ID), meta, stringsAsFactors = F)
# 
# # set column to be used for merging in meta_data
# meta_key <- as.character(unique(meta$Plate_ID))

# replace entries in data$Plate_ID to match meta$Plate_ID
# uses meta_merge and data_merge as keys
# can do this loop as sapply as well
# merge the list to include the key pairs that will be matched up to be passed on the replacement function
# 
# for (i in 1:length(meta_key)) { raw$join_Plate_ID[raw$join_Plate_ID==raw_key[i]] <- meta_key[i] }

# join by Plate_ID column and Well_Name
complete <- full_join(meta, raw, by = c("Scan_ID" = "Plate_ID", "Well_Name" = "Well_Name"))

# fully assembled dataset
datatable(complete)

# write fully assembled data
write_csv(complete, paste0(out_dir_data, out_data, ".csv"))




```
