---
title: "Raw/Meta-Data Merge"
subtitle: "Use with discretion and QC the results!"
author: "Mario Niepel"
date: "December 2020"
version: "0.4"
output:
     html_document:
          toc: true
          toc_depth: 2
          theme: united 
code_folding: hide
          
---

# To Do

- generate code outline with chunk names and hash tags for indentations
- function to plot 96/384 well plate based on well_level_ids
     - lapply to run through list of well_level_IDs
     - print 1 graphical outline per 'is_unique' on a per-plate basis
- html output for summary
- set up QC plots as plot.ly (using ggplotly)
- harmonize saving as TSV vs. CSV

1) clean up and comment current code -- include (import/output steps into markdown)
3) fortify meta-data assembly (find and remove rows of 1-24 and columns of a-p)
4) fortify raw/meta merger
5) automate initial cleanup/plotting of merged data
6) function-ize current scripts -- clear spots for user input/edits

QC plots
- add shapes for 'Drug_ID overlayed onto color' instead for written name?
- reorder character legend with NULL as the last element

- add summary table for raw data once appended

- put is more safety checks to test if data input is as expected
	- eg does data in raw match up to data in meta
	- nrow(raw) must be a multiple of nrow(meta)

```{r loading libraries, include=FALSE}

library(tidyverse)
library(patchwork)
library(DT)

```




# Summary and requirements:


This script takes a CSV file from Excel or another similar program containing metadata of experiments and converts it semi-automatically to a long dataframe suitable to receive output from virtually any other instrument that puts the data on a per-plate basis either into a single table or a row/column format. This script only deals with the metadata. 'raw_data_input' deals with the experimental results. A third script 'meta_raw_merge' combines the two and performs some basic QC. The thought is that the output from that script is ready for deeper analysis, plotting of results, extracting data for other experiments, ...

## MetaData requirements:

The file contains two parts: Plate-level data first and Well-level data second. For now only 384 well plates are supported, but the script should be easily adjustable for 96 well plates or any other plate-based format. Spaces in any headers will be replaced with underscores

## Plate-level data:

- A table with columns for variables/attributes with one row for each physical plate used in the experiment.
- **One column needs to be labeled 'Experiment_ID' and must contain an identical ID in each row.** Best practices take the date of 'day0' of the experiment and append your initials. For an experiment I started on April 16th 2019 the ID would be "20190416_MN"
- **One column needs to be labeled 'Plate_ID' and must contain a unique ID in each row that should not overlap with any other experiments** in case results will letter be merged. Best practices are to simply take the experiment ID and append _01, _02, ...This way there will be no overlap. Each entry has to be a complete column and contain a header and one entry for each plate. Missing values will create ambiguities when addressing the data later and can not be tolerated


## Well-level data:

- each well-level treatment needs a complete 'block' of information made of two parts. Variable information for the type of treatment and qualitative or quantitative information about the treatment for each well
- the type of treatment has to be a block of four entries:

+-------------+------------+
|  variable   | type       |
+-------------+------------+
|  descriptor | attribute  |
+-------------+------------+

"variable" needs to be entered for the script to recognize a new column
type needs to be entered as "numeric" or "factor" to designate if there is a continuous variable or not
descriptor denotes a treatment class that makes sense (e.g. drug, cytokine, media, gRNA, cell_no, ...)
if a given descriptor has multiple attributes they are combined with the descriptors. For all descriptors the most basic attribute should be ID. Additional ones for a drug could be: concentration, the vehicle, and treatment time. This would give drug_ID, drug_conc, drug_vehicle, and drug_treattime. And each one of these variables would get their own block of row/column entries.

- the qualitative/quantitative block for each variable needs to be regular and complete (no empty cells in 16 rows x 24 columns). Empty cells can be filled either with NA, NULL, NaN, or whatever other label makes sense to you. The dataframe will contain these entries as it is being assembled, but they can easily be filtered out later. 

## Different plate layouts for the same treatment

In some cases the well-level treatemnts differ from plate to plate. In this case the annotation has to be in both the plate- and well-level. In the Plate-Level column use different names to indicate the different plate layouts: e.g. (Drug_ID / group1, group2, group1, group2). At the well level label things as usual and simply include duplicate variables for the two different layouts (Drug_ID). The script will automatically assign the first group at the plate level to the first well level layout of the corresponding ID. There have to be the same number of layouts as groups and of course names have to be exactly identical.

```{r designation of files and directory}

# Import of meta data and organization of folders and files

# user input of which meta file to load
# format has to be according to specifications laid out
meta_in_file <- "20201130_AL_metadata"

# user input of which raw data files to load
raw_in_file <- c("20201130_AL_IncuCyte_parsed_Plate2.tsv",
                 "20201130_AL_IncuCyte_parsed_Plate1.tsv",
                 "20201130_AL_PlateReader_parsed_Plate1.tsv",
                 "20201130_AL_PlateReader_parsed_Plate2.tsv")

# name of assembled complete data file
out_data <- paste0("20201130_AL_data", "_complete")
out_qc <- paste0("20201130_AL_data", "_QC")

# input and output directories 
in_dir <- "./data/"
out_dir_data <- "./data/"
out_dir_plots <- "./data/"

# match up of meta and raw data
join_by <- c(
     # Plate column in meta = raw
     "Plate_ID" = "Plate_ID", 
     # Well column in meta = raw
     "Well_ID" = "Well_ID")

```




```{r file input & QC, include=FALSE}

# check if out_dirs exists; if not make it
if (!dir.exists(out_dir_data)) {dir.create(out_dir_data)}
if (!dir.exists(out_dir_plots)) {dir.create(out_dir_plots)}

#read input_file
suppressMessages(meta <- read_csv(paste0(in_dir, meta_in_file, ".csv"), col_names = FALSE))

# remove all NA rows and columns
meta <- meta[rowSums(is.na(meta)) != ncol(meta),]
meta <- meta[,colSums(is.na(meta)) != nrow(meta)]

# add filler into empty cell in 1/1 if empty

# Validate the presence of Experiment_ID and Plate_ID
if (length(which(meta=="Experiment_ID")) == 0) {
     stop("Experiment_ID not found. Please fix input.")
}

if (length(which(meta=="Experiment_ID")) > 2) {
     stop("Experiment_ID definition is not unique. Please fix input.")
}

if (length(which(meta=="Plate_ID")) == 0) {
     stop("Plate_ID not found. Please fix input.")
}


```




```{r split plates/wells, echo=FALSE, warning=TRUE}

# identify entry with 'Experiment_ID'
row <- which(meta=="Experiment_ID", arr.ind = T)[1]
col <- which(meta=="Experiment_ID", arr.ind = T)[2]

# set name of Experiment_ID
Experiment_ID <- toString(meta[row+1,col])

# subset all rows of meta that contain "experiment ID" 
# %in% checks the resulting vectorized evaluation of "(meta[,col] == "Experiment_ID" | meta[,col] == Experiment_ID)" and tests if it is TRUE
plates <- meta[(meta[,col] == "Experiment_ID" | meta[,col] == Experiment_ID) %in% TRUE, ]

# subset all rows of meta with the logical inverse of plates
wells <- meta[!((meta[,col] == "Experiment_ID" | meta[,col] == Experiment_ID) %in% TRUE),]

rm(meta)

```


```{r clean up and QC, echo=FALSE, warning=TRUE}
# clean up and sanity check for plates 

# remove NA columns as above
plates <- plates[,colSums(is.na(plates)) != nrow(plates)]

# check for presence of any NA in rows
if (sum(rowSums(is.na(plates))) > 0) {
     stop("Plate annotation is incomplete. Missing values can't be interpreted.")
}

# convert first row into names and remove that row
colnames(plates) <- as.character(unlist(plates[1,]))
plates <- plates[-1,]

# Set up Experiment_ID level and Plate_ID

# check that there is only one entry in Experiment_ID
if (n_distinct(plates$Experiment_ID)!=1) {
     stop("Experiment_ID is not identical across all plates.")
}

# set up experiment_level dataframe and remove from plates
exp_level <- data.frame(Experiment_ID=plates$Experiment_ID, stringsAsFactors = F)
plates <- plates[ , !(names(plates)=="Experiment_ID")]

# check that there are nrow entries in Experiment_ID
if (n_distinct(plates$Plate_ID)!=nrow(plates)) {
     stop("Plate_ID is not unique across all plates.")
}

# set up plate_level dataframe and remove from plates
plate_level <- data.frame(Plate_ID=plates$Plate_ID, stringsAsFactors = F) 
plates <- plates[ , !(names(plates)=="Plate_ID")]

```


```{r parse plate section with plate/well specs, echo=FALSE, warning=TRUE}

# number of unique vals per column
unique_vals <- unname( sapply(plates, function(x) length(unique(x))) )  
unique_cols <- unname( sapply(plates, function(x) length(unique(x))==1) ) 

# append all unique columns to exp_level
exp_level <- cbind(exp_level, plates[ ,unique_vals==1])

# append the opposite selection to plate_level
plate_level <- cbind(plate_level, plates[ ,unique_vals!=1])

# create and insert Plate_No column
plate_level$Plate_No <- as.integer(1:nrow(plates)) 
plate_level <- plate_level[ , c(1, ncol(plate_level), 2:(ncol(plate_level)-1))]

# Generate Plate_Row, _Col, Well_No
well_level <- data.frame(Plate_Row = (rep(LETTERS[seq(1:16)],24)),
                         Plate_Col = as.factor(trunc((seq(1:384)+15)/16)),
                         Well_No = as.factor(seq(1:384)))

# add column for Well_ID (A01-P24)
well_level <- well_level %>% 
     # Well_ID is concatenation of Plate_Row and Plate_Col
     # str_pad adds a leading character ("0") to width (2)
     mutate(Well_ID = as.factor(paste0(Plate_Row, str_pad(Plate_Col, 2, pad = "0")))) %>%
     # sort by well number
     arrange(Well_No)



```


```{r extract well level annotation, echo=FALSE, warning=TRUE}

######################################################################
### could write this as mapply as well or lapply with string/split
### variables is a data frame
### turn into a list that contains x/y pairs
### pass on this list to sapply to slice out and replace
######################################################################

# identify positions of "variable"
variables <- data.frame(which(wells=="variable", arr.ind=TRUE))       # identify x/y of 'variables' 
variables <- variables[order(variables$row),]                         # sort variables table by row

# join 384 row well_level info with readouts
end_row <- 0
for (i in 1:nrow(variables)) {                                   

# empty out dataframe
     slice <- data.frame()
          
# put row/col value of current plate map in row and col
     row <- variables[i,"row"]
     col <- variables[i,"col"]

# identify variable type, names, and attribute
     type <- wells[row, col+1]
     name <- wells[row+1, col]
     attribute <- wells[row+1, col+1]

# slice out plate based on location of 'variables'
     if  ( i==nrow(variables) ) ( end_row <- nrow(wells) ) else ( end_row <- variables[i+1,"row"]-1 )
     slice <- wells %>% slice((row+2):end_row)

# remove NA rows and NA columns
     slice <- slice[rowSums(is.na(slice)) != ncol(slice),] # remove cols of all NA
     slice <- slice[,colSums(is.na(slice)) != nrow(slice)] # remove rows of all NA     

# values are the readout of the parsed plate as single vector
     values <- unlist(slice, use.names = F)
 
# append column to well_level dataframe
     if (type=="numeric") {
               well_level <- cbind(well_level, as.numeric(values))
          } else {
               well_level <- cbind(well_level, values, stringsAsFactors = F)
          }

     # add proper metadata heading to new df column 'Name_Attribute' e.g drug_id
     names(well_level)[ncol(well_level)] <- paste0(name,"_",attribute)
     
# relevel so that NULL is last element
     n <- ncol(well_level)
     if (any(levels(well_level[ , n]) == "NULL")) 
      { well_level[ , n] <- factor(well_level[ , n] , levels = c(setdiff(levels(well_level[ , n]), "NULL"),"NULL")) }                                       

}                                                                   


# clean up variables
rm(col, end_row, i, n, row, values, unique_cols, unique_vals, type, slice, name, attribute, variables)

```




```{r assemble plate and well data, echo=FALSE, warning=TRUE}

# assemble the full data frame from plate and well info

# extract level variables
well_level_names <- c(colnames(well_level[5:ncol(well_level)]))
plate_level_names <- c(colnames(plate_level))
exp_level_names <- c(colnames(exp_level)) 

# join experiment and plate-level metadata
# Duplicate each row in plates 384x
# lastly reorder the file by plate No 
exp_plate_level <- cbind(exp_level, plate_level) %>% slice(rep(row_number(), 384)) %>% arrange(Plate_No)

# combine wells_long and plates_long
meta_processed <- cbind(exp_plate_level, well_level)

# check if there are duplicates in plate level and well level that need special attention
duplicate <- intersect(plate_level_names, well_level_names)

# if there are no duplicates skip this module
 if (length(duplicate) != 0) {

     # set up list to collect the vectors for merge
     master <- vector(mode = "list", length(duplicate))
     names(master) <- duplicate

     # cycle through each duplicate
     for (i in 1:length(duplicate)) {

          # set up list of vectors with what to merge
          merge_cols_well_level <- str_which(colnames(well_level), duplicate[i])
          merge_cols_plate_level <- str_which(colnames(plate_level), duplicate[i])

          # cycle through each plate for each duplicate
          for (j in 1:nrow(plate_level)) {

               # which level is the current duplicate to merge
               level <- min( which (plate_level[,merge_cols_plate_level] == plate_level[j,merge_cols_plate_level] ) )

               # find which elements to add and append them to the merge list
               master[[i]] <- append(master[[i]], well_level[1:384,merge_cols_well_level[level]])

          } # end of j-loop

     } # end of i-loop

# remove relevant names from plate level
plate_level[colnames(plate_level) %in% duplicate] <- NULL
plate_level_names <- c(colnames(plate_level))

# remove relevant names from well level
well_level[colnames(well_level) %in% duplicate] <- NULL
well_level_names <- c(colnames(well_level), duplicate)

# remove duplicate columns from meta_processed
meta_processed[ colnames(meta_processed) %in% duplicate ] <- NULL

# merge remaining columns in meta_processed with master list
meta_processed <- cbind( meta_processed, data.frame(master, stringsAsFactors = F))

# remove variables used for loop
rm(master, merge_cols_well_level, merge_cols_plate_level, level, duplicate, i, j)

} 

# sort by plate and well number
meta_processed <- arrange(meta_processed, Plate_No, Plate_Col)

# give proper row names to complete file
rownames(meta_processed) <- c(seq(1:nrow(meta_processed)))

# rearrange columns by using relocate
# go via creation of 'new order' vector in case exp_level or plate_level is empty
# this causes dplyr to throw an error
new_order <- c("Experiment_ID", exp_level_names, "Plate_ID", "Plate_No", "Plate_Row", "Plate_Col", plate_level_names, "Well_ID", "Well_No") 
meta_processed <- meta_processed %>% relocate (new_order[]) 
rm(new_order)

# save processed data
write_csv(meta_processed,paste0(out_dir_data, meta_in_file, "_formatted.csv"))


```




## MetaData Tables

```{r data summary, echo=FALSE}

# table representation of experiment level annotation

# select only columns with names in 'exp_level_names'
exp_level_table <- meta_processed[1,exp_level_names]
# extract values w/o names as characters from every column
exp_level_values <- list(values=unlist(lapply(exp_level_table, as.character), use.names = FALSE))
# generate new df with name and value column
exp_level_table <- data.frame(names=exp_level_names, values = exp_level_values )
# experiment_level table
datatable(exp_level_table)

# table representation of plate level annotation

# select all columns based on plate level name and extract distinct values based on each Plate ID
plate_level_table <- meta_processed %>% select(all_of(plate_level_names)) %>% group_by(Plate_ID) %>% distinct()
# plate_level table
datatable(plate_level_table)

# table representation of well level annotation

# grepl("_ID", well_level) Searches for any list entry that contains the string "_ID" in the list of "well_level"
# substr cutting of the last three characters "_ID" through substring(x, 1, nchar(x)-3)
well_level_ids <- substr(well_level_names[grepl("_ID", well_level_names)], 1, nchar(well_level_names[grepl("_ID", well_level_names)])-3)
# remove Well_ID and Well_No
well_level_ids <- well_level_ids[well_level_ids != "Well"]

# subset all columns that start with well_level_ids
well_level_table <- meta_processed %>% select(all_of(starts_with(well_level_ids))) %>% lapply(unique)
# convert into characters so table can be better manipulated
well_level_table <- lapply(well_level_table, as.character)
# fill up empties with NA and turn into a dataframe
# lapply makes length of all vectors the same as the max which fills them up with NA
well_level_table <- data.frame(lapply(well_level_table, function(x) { length(x) <- max(lengths(well_level_table)); x }))
# well_level table
datatable(well_level_table)

# full metadata
datatable(meta_processed, options = list(
  columnDefs = list(list(className = 'dt-center', targets = 1:ncol(meta_processed))),
  pageLength = 16,
  lengthMenu = c(8, 16, 96, 384), class = "stripe hover compact",
  style="width:100%",
  scrollX = T)
)



```



## MetaData QC Figures

```{r visual data summary, echo=FALSE, fig.width=16, fig.height=12 }

####################################################################################

qc_plot <- function(col, plate, df) { 

if ( is.numeric( df[[col]] ) ) {
     
     nudge <- c( rep(0.25, times=16), rep(-0.25, times=16))    
    
     label <- col
     if (str_sub(label, start= -3) != "_ID" ) { label <- paste0(unlist(strsplit(label, "_"))[[1]], "_ID")}

     df %>% filter(Plate_ID == plate) %>% 

     ggplot(aes(x = Plate_Col, y = Plate_Row)) + 
     coord_fixed() + 
     scale_x_discrete(limits = c(as.character(seq(1, 24)))) +
     scale_y_discrete(limits = rev(levels(df$Plate_Row))) +
     theme(legend.position="bottom", legend.box = "horizontal", text = element_text(size = 20)) +
     
     labs(title = paste("Readout:", col, "\n1st unique plate:", plate) ,
     caption = paste("Experiment ID:", Experiment_ID)) +
     
     geom_tile(aes(fill = .data[[col]]), color = "black") +
     scale_fill_viridis_c(begin = 0, alpha = 0.5, direction = -1, na.value = "gray",
                          trans = "log", breaks = c(10, 3.1, 1, 0.31, 0.1, 0.03, 0.01, 0.003, 0.001)) +
     guides(fill = guide_colourbar(barwidth = unit(6, "in"), barheight = unit(.6, "in"))) +
     geom_text(aes(x = Plate_Col, y = Plate_Row, label = .data[[label]]), check_overlap = TRUE, nudge_y = nudge, size = 4) 

     } else {
     
     nudge <- c( rep(0.25, times=16), rep(-0.25, times=16))    
    
     label <- col
     if (str_sub(label, start= -3) != "_ID" ) { label <- paste0(unlist(strsplit(label, "_"))[[1]], "_ID")}

     df %>% filter(Plate_ID == plate) %>% 

     ggplot(aes(x = Plate_Col, y = Plate_Row)) + 
     coord_fixed() + 
     scale_x_discrete(limits = c(as.character(seq(1, 24)))) +
     scale_y_discrete(limits = rev(levels(df$Plate_Row))) +
     theme(legend.position="bottom", legend.box = "horizontal", text = element_text(size = 20)) +
     
     labs(title = paste("Readout:", col, "\n1st unique plate:", plate) ,
     caption = paste("Experiment ID:", Experiment_ID)) +
     
     geom_tile(aes(fill = .data[[col]]), color = "black", size = 0.5) +
     scale_fill_viridis_d(begin = 0.2, alpha = 0.5, direction = -1, na.value = "gray") +
     geom_text(aes(x = Plate_Col, y = Plate_Row, label = .data[[label]]), check_overlap = TRUE, nudge_y = nudge, size = 4) 
     
     }

}


####################################################################################


# generate lookup table of unique plates that need to be summarized
summary_plates <- 
     meta_processed %>%
     group_by(Plate_ID) %>%
     summarize(across(all_of(well_level_names[5:length(well_level_names)]), paste, collapse = ""), .groups = "drop") %>%
     mutate(across(all_of(well_level_names[5:length(well_level_names)]), ~!duplicated(.)))


# generate list with entries for each column listing plates to plot
list_summary_plates <- sapply(summary_plates[-1], function(x) { summary_plates$Plate_ID[x] })

# generate sum_plate and sum_ID vectors for summary plots
sum_IDs <- c(rep(names(list_summary_plates), lengths(list_summary_plates)))
sum_plates <- as.character(unlist(list_summary_plates, use.names = F))

# call qc_plot with mapply using the IDs and corresponding plate as list arguments
qc_plots <- mapply(qc_plot, sum_IDs, sum_plates, MoreArgs = list(meta_processed), SIMPLIFY = FALSE)

# print off all QC plots
# maybe set up as plot.ly (using ggplotly)
for (p in 1:length(qc_plots)) { print(qc_plots[[p]]) }

# Export to PDF
#   
# pdf(file = paste0(out_dir, out_file, ".pdf"),   # The directory you want to save the file in
#     width = 16, # The width of the plot in inches
#     height = 12) # The height of the plot in inches
# 
# for (i in 1:4) {
#      print(composite1[[i]])
#      print(composite2[[i]])
# }
# dev.off()
# invisible(NULL)

```




```{r raw data import and assembly -- adjusted, warning=TRUE, include=FALSE}

# grab meta from meta_processed
meta <- meta_processed

# read all data files into a list using sapply
raw_list <- lapply(paste0(in_dir, raw_in_file), function(x) read_tsv(x, col_names = TRUE))

# convert names into collapsed strings
raw_names <- sapply(raw_list, function(x) str_c(names(x), collapse = ""))

# generate index of which files have identical columns
index <- unname(sapply (raw_names, function (x) which(unique(raw_names) == x) ))

# list for full_join
merge_list <- vector(mode = "list", length = length(unique(raw_names)))

# loop over number of unique names to merge by identical files by rbind
for (i in 1:length(unique(raw_names))) {
	
	# remove all list items that are not equivalent to index i
	temp <- raw_list
	temp[which(index != i)] <- NULL
	# fill list[[i]] with first dataframe to merge
	merge_list[[i]] <- temp[[1]]
	temp[[1]] <- NULL
	
	# rbind remaining items in temp with merge_list by rbind
	for (j in 1:length(temp) ) { merge_list[[i]] <- rbind ( merge_list[[i]], temp[[j]] ) } 
}

# see if joining is needed
if (length(merge_list) > 1) {
	# plate 1st item into final raw dataset
	raw <- merge_list[[1]]
	
	# loop over remaining merge_list
	for (i in 2:length(merge_list)) {
		# perform full join by intersection of col names
		raw <- full_join(raw, merge_list[[i]], by = join_by ) 
	}
}

# clean-up
rm(i, j, temp, merge_list, raw_list, raw_names, index)

# identify columns that are overlapping between meta and raw
duplicate <- intersect(names(meta), names(raw))

# remove duplicate columns used for joining
duplicate <- setdiff(duplicate, unlist(join_by))

# tag all remaining columns with '_raw'
names(raw)[names(raw) %in% duplicate] <- paste0(names(raw)[names(raw) %in% duplicate], "_raw")


```



```{r raw data import and assembly, warning=TRUE, include=FALSE}

# # grab meta from meta_processed
# meta <- meta_processed
# 
# # read all data files into a list using sapply
# raw <- lapply(paste0(in_dir, raw_in_file), function(x) read_tsv(x, col_names = TRUE))
# 
# # reduce left iterates through a list with many items and consecutively applies the function to pairs in the list
# # that's how reduce works to cycle through columns when checking for NA, but not through rows.
# raw <- raw %>% reduce(full_join, by = join_by)
# 
# # identify columns that are overlapping between meta and raw
# duplicate <- intersect(names(meta), names(raw))
# 
# # remove duplicate columns used for joining
# duplicate <- setdiff(duplicate, unlist(join_by))
# 
# # tag all remaining columns with '_raw'
# names(raw)[names(raw) %in% duplicate] <- paste0(names(raw)[names(raw) %in% duplicate], "_raw")


```

## Assembled DataSet

```{r clean up & merge with meta, echo=FALSE, warning=TRUE}

# Clean up data to be merged

# Clean up column names
colnames(raw)[which(colnames(raw)=="% Positive W2 (MultiWaveScoring)")] <- "Fraction_W2"
colnames(raw)[which(colnames(raw)=="Cell: W2 Nucleus Average Intensity (MultiWaveScoring)")] <- "Nuc_W2"
colnames(raw)[which(colnames(raw)=="Cell: W2 Cytoplasm Average Intensity (MultiWaveScoring)")] <- "Cyto_W2"
colnames(raw)[which(colnames(raw)=="Cell: W2 Cell Average Intensity (MultiWaveScoring)")] <- "Cell_W2"
colnames(raw)[which(colnames(raw)=="% Positive W3 (MultiWaveScoring)")] <- "Fraction_W3"
colnames(raw)[which(colnames(raw)=="Cell: W3 Nucleus Average Intensity (MultiWaveScoring)")] <- "Nuc_W3"
colnames(raw)[which(colnames(raw)=="Cell: W3 Cytoplasm Average Intensity (MultiWaveScoring)")] <- "Cyto_W3"
colnames(raw)[which(colnames(raw)=="Cell: W3 Cell Average Intensity (MultiWaveScoring)")] <- "Cell_W3"
colnames(raw)[which(colnames(raw)=="Total Cells (MultiWaveScoring)")] <- "Cell_No"
colnames(raw) <- gsub(" ", "_", colnames(raw))

# remove all rows where Plate_ID is NA
raw <- raw[ !is.na(raw$Plate_ID)  , ]

# join by Plate_ID column and Well_ID
complete <- full_join(meta, raw, by = join_by)

# fully assembled dataset
datatable(complete, options = list(
  columnDefs = list(list(className = 'dt-center', targets = 1:ncol(meta_processed))),
  pageLength = 16,
  lengthMenu = c(8, 16, 96, 384), class = "stripe hover compact",
  style="width:100%",
  scrollX = T)
)

# write fully assembled data
write_csv(complete, paste0(out_dir_data, out_data, ".csv"))




```

### Version History


## version 0.5


##version 0.4
# implemented the input and QC of plates that are treated with different layouts of the same treatments
	- e.g different cell lines on separate plates, but arranged differently 
	- two drug layouts
	- columns between plate-level and well level have to agree
  
# switched to exclusively non-factor columns for anything that is not a fixed designation such as Exp ID, and physical descriptors of the plate
# tagged columns in raw section that are identical in meta. Meta columns remain primary and raw duplicate columns are marked
# cleaned up code 
# updated multi-file raw data merger
	- implemented that any number of raw data files can be used
	- files with identical columns will be first joined by rbind
	- all remaining files with be joined by full_join

 
## version 0.3
# generated all new graphic plate QC and summary tables
# identifies 'unique' plates and visualizes them
